---
title: "Project-Hawks"
author: "Julia Cuellar"
date: '2021-7-4'
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## EDA
```{r eval = TRUE, echo = FALSE, warning = FALSE, message = FALSE}
library(knitr)
library(latexpdf)
library(tinytex)
library(dplyr)
library(tidyverse)
library(cluster)
library(hrbrthemes)
library(NbClust)
library(factoextra)
library(ggplot2)
setwd("C:/Users/jcuel/Videos/Bellevue/Summer 2021/Professor Williams/Projects/Project 1")
hawk_df = read.csv("Hawks.csv")
```

**Hawks Data:**
```{r eval = TRUE, echo = FALSE}
head(hawk_df)
```

**Summary of Hawks Data:**
```{r eval = TRUE, echo = FALSE}
summary(hawk_df)
```

```{r eval = TRUE, echo = FALSE}
hawk_df = hawk_df %>% select(-X)
hawk_df = hawk_df[ , names(hawk_df) %in% c("Month", "Year", "CaptureTime", "BandNumber", "Species", "Age", "Wing", "Weight", "Culmen", "Hallux", "Tail", "StandardTail", "Tarsus")]
```

**Display nulls of Hawks Data:**
```{r eval = TRUE, echo = FALSE}
is.null(hawk_df)
```

**Display NA's of Hawks Data:**
```{r eval = TRUE, echo = FALSE}
sum(is.na(hawk_df$Month))
sum(is.na(hawk_df$CaptureTime))
sum(is.na(hawk_df$Species))
sum(is.na(hawk_df$Age))
sum(is.na(hawk_df$Hallux))
```

**Redisplay NA's of Hallux:**
```{r eval = TRUE, echo = FALSE}
hawk_df = replace_na(hawk_df, list(`Hallux` = mean(hawk_df$'Hallux', na.rm = TRUE)))
sum(is.na(hawk_df$Hallux))
```

**Redisplay Hawks Data:**
```{r eval = TRUE, echo = FALSE}
head(hawk_df)
```

**Project Month plot:**

```{r eval = TRUE, echo = FALSE}
month_plot = hist(hawk_df$Month, main = 'Month', xlab = 'month', ylab = "Frequency", col = c("yellow", "orange", "brown", "red"))
```

**Project Species plot:**

```{r eval = TRUE, echo = FALSE}
cp_plot = ggplot(hawk_df, aes(x = Month, y = CaptureTime)) + geom_line() + labs(title = "Capture Time")
species_plot = barplot(table(hawk_df$Species), main = "Species", xlab = "species", ylab = "Count", col = c("orange", "red", "brown"), legend = c("CH", "RT", "SS"))
text(species_plot, 0, round(table(hawk_df$Species), 1), cex = 1, pos = 3)
```

**Project Age plot:**

```{r eval = TRUE, echo = FALSE}
age_plot = barplot(table(hawk_df$Age), main = "Age", xlab = "age", ylab = "Count", col = c("pink", "blue"), legend = c("A", "I"))
text(age_plot, 0, round(table(hawk_df$Age), 1), cex = 1, pos = 3)
```

**Project Hallux plot:**

```{r eval = TRUE, echo = FALSE}
boxplot(hawk_df$Hallux, main = 'Hallux', xlab = 'hallux', col = 'purple')
```


## Cluster Model of the Species of Hawks for Age
```{r eval = TRUE, echo = FALSE}
library(dplyr)
library(tidyverse)
setwd("C:/Users/jcuel/Videos/Bellevue/Summer 2021/Professor Williams/Projects/Project 1")
hawk_df = read.csv("Hawks.csv")
hawk_df = hawk_df %>% select(-X)
hawk_df = hawk_df[ , names(hawk_df) %in% c("Month", "Year", "CaptureTime", "BandNumber", "Species", "Age", "Wing", "Weight", "Culmen", "Hallux", "Tail", "StandardTail", "Tarsus")]
hawk_df = replace_na(hawk_df, list(`Hallux` = mean(hawk_df$'Hallux', na.rm = TRUE)))
```

**Cross table of Species of Hawks vs Age:**

```{r eval = TRUE, echo = FALSE}
adult = subset(hawk_df, Age == "A")
adult_ch = subset(adult, Species == "CH")
adult_rt = subset(adult, Species == "RT")
adult_ss = subset(adult, Species == "SS")
imma = subset(hawk_df, Age == "I")
imma_ch = subset(imma, Species == "CH")
imma_rt = subset(imma, Species == "RT")
imma_ss = subset(imma, Species == "SS")
```

Species       | Adult | Immature | **Total**
------------- | ----- | -------- | -------
Cooper's      |  32   |  38      |  70
Red-tailed    |  123  |  454     |  577
Sharp-shinned |  69   |  192     |  261
------------- | ----- | -------- | -------
**Total**     |  224  | 684      | 908

## Cluster Model of the Species of Hawks for Month
```{r eval = TRUE, echo = FALSE, warning = FALSE}
library(dplyr)
library(tidyverse)
setwd("C:/Users/jcuel/Videos/Bellevue/Summer 2021/Professor Williams/Projects/Project 1")
hawk_df = read.csv("Hawks.csv")
hawk_df = hawk_df %>% select(-X)
hawk_df = hawk_df[ , names(hawk_df) %in% c("Month", "Year", "CaptureTime", "BandNumber", "Species", "Age", "Wing", "Weight", "Culmen", "Hallux", "Tail", "StandardTail", "Tarsus")]
hawk_df = replace_na(hawk_df, list(`Hallux` = mean(hawk_df$'Hallux', na.rm = TRUE)))
month_df = hawk_df$Month
month_df = data.frame(month_df)
```

**Exhibit Dissimilarity Matrix for Month:**

```{r eval = TRUE, echo = FALSE}
library(cluster)
library(hrbrthemes)
library(NbClust)
library(factoextra)
dist_month = daisy(month_df, metric = "gower")
month_gradient_color = list(low = "chocolate1",  high = "white")
fviz_dist(dist_month, gradient = month_gradient_color, order = T)
```

**Display # of clusters for Month:**

```{r eval = TRUE, echo = FALSE, warning = FALSE}
num_clust_month = NbClust(diss = dist_month, distance = NULL, min.nc = 2, max.nc = 5, method = "median", index = "silhouette")
num_clust_month$Best.nc
```

**Project MDS plot for Month:**

```{r eval = TRUE, echo = FALSE, warning = FALSE}
library(ggplot2)
month_pam = pam(dist_month, 4)
month_mds = as.data.frame(cmdscale(dist_month, 2))
month_mds$num_clust_month = as.factor(month_pam$clustering)
ggplot(month_mds, aes(x = V1, y = V2, color = num_clust_month)) + geom_point() + theme_ipsum() + labs(title = "MDS plot for Month", subtitle = "Colored by PAM cluster") + scale_color_brewer(palette = "Set1")
```

**Project Species & PAM Cluster for Month:**

```{r eval = TRUE, echo = FALSE, warning = FALSE}
hawk_df$Species = as.factor(month_mds$num_clust_month)
ggplot(hawk_df, aes(x = Species)) + geom_bar(stat = "count", position = "dodge", width = 0.7) + theme_ipsum(grid = "Y") + scale_fill_brewer(palette = "Set2") + labs(title = "Species and PAM cluster", subtitle = "for Month") + xlab("Month") + ylab("count")
```

**Cross table of Species of Hawks vs Month:**
```{r eval = TRUE, echo = FALSE}
library(dplyr)
library(tidyverse)
setwd("C:/Users/jcuel/Videos/Bellevue/Summer 2021/Professor Williams/Projects/Project 1")
hawk_df = read.csv("Hawks.csv")
hawk_df = hawk_df %>% select(-X)
hawk_df = hawk_df[ , names(hawk_df) %in% c("Month", "Year", "CaptureTime", "BandNumber", "Species", "Age", "Wing", "Weight", "Culmen", "Hallux", "Tail", "StandardTail", "Tarsus")]
hawk_df = replace_na(hawk_df, list(`Hallux` = mean(hawk_df$'Hallux', na.rm = TRUE)))
aug = subset(hawk_df, Month == 8)
aug_ch = subset(aug, Species == "CH")
aug_rt = subset(aug, Species == "RT")
aug_ss = subset(aug, Species == "SS")
sept = subset(hawk_df, Month == 9)
sept_ch = subset(sept, Species == "CH")
sept_rt = subset(sept, Species == "RT")
sept_ss = subset(sept, Species == "SS")
oct = subset(hawk_df, Month == 10)
oct_ch = subset(oct, Species == "CH")
oct_rt = subset(oct, Species == "RT")
oct_ss = subset(oct, Species == "SS")
nov = subset(hawk_df, Month == 11)
nov_ch = subset(nov, Species == "CH")
nov_rt = subset(nov, Species == "RT")
nov_ss = subset(nov, Species == "SS")
```

Species       | August | September | October | November | **Total**
------------- | ------ | --------- | ------- | -------- | --------
Cooper's      |  0     |  31       |  34     |  5       | 70
Red-tailed    |  1     |  150      |  314    |  112     | 577
Sharp-shinned |  0     |  112      |  114    |  35      | 261
------------- | ------ | --------- | ------- | -------- | --------
**Total**     |  1     |  293      |  462    |  152     | 908

## Cluster Model of the Species of Hawks for Hallux
```{r eval = TRUE, echo = FALSE}
library(dplyr)
library(tidyverse)
setwd("C:/Users/jcuel/Videos/Bellevue/Summer 2021/Professor Williams/Projects/Project 1")
hawk_df = read.csv("Hawks.csv")
hawk_df = hawk_df %>% select(-X)
hawk_df = hawk_df[ , names(hawk_df) %in% c("Month", "Year", "CaptureTime", "BandNumber", "Species", "Age", "Wing", "Weight", "Culmen", "Hallux", "Tail", "StandardTail", "Tarsus")]
hawk_df = replace_na(hawk_df, list(`Hallux` = mean(hawk_df$'Hallux', na.rm = TRUE)))
hallux_df = hawk_df$Hallux
hallux_df = data.frame(hallux_df)
```

**Exhibit Dissimilarity Matrix for Hallux:**

```{r eval = TRUE, echo = FALSE}
library(cluster)
library(hrbrthemes)
library(NbClust)
library(factoextra)
dist_hallux = daisy(hallux_df, metric = "gower")
hallux_gradient_color = list(low = "thistle2",  high = "white")
fviz_dist(dist_hallux, gradient = hallux_gradient_color, order = T)
```

**Display # of clusters for Hallux:**

```{r eval = TRUE, echo = FALSE, warning = FALSE}
num_clust_hallux = NbClust(diss = dist_hallux, distance = NULL, min.nc = 2, max.nc = 15, method = "median", index = "silhouette")
num_clust_hallux$Best.nc
```

**Project MDS plot for Hallux:**

```{r eval = TRUE, echo = FALSE, warning = FALSE}
library(ggplot2)
hallux_pam = pam(dist_hallux, 2)
hallux_mds = as.data.frame(cmdscale(dist_hallux, 2))
hallux_mds$num_clust_hallux = as.factor(hallux_pam$clustering)
ggplot(hallux_mds, aes(x = V1, y = V2, color = num_clust_hallux)) + geom_point() + theme_ipsum() + labs(title = "MDS plot for Hallux", subtitle = "Colored by PAM cluster") + scale_color_brewer(palette = "Set1")
```

**Project Species & PAM Cluster for Hallux:**

```{r eval = TRUE, echo = FALSE, warning = FALSE}
hawk_df$Species = as.factor(hallux_mds$num_clust_hallux)
ggplot(hawk_df, aes(x = Species)) + geom_bar(stat = "count", position = "dodge", width = 0.7) + theme_ipsum(grid = "Y") + scale_fill_brewer(palette = "Set2") + labs(title = "Species and PAM cluster", subtitle = "for Hallux") + xlab("Hallux") + ylab("count")
```

**Cross table of Species of Hawks vs Hallux:**
```{r eval = TRUE, echo = FALSE}
library(dplyr)
library(tidyverse)
setwd("C:/Users/jcuel/Videos/Bellevue/Summer 2021/Professor Williams/Projects/Project 1")
hawk_df = read.csv("Hawks.csv")
hawk_df = hawk_df %>% select(-X)
hawk_df = hawk_df[ , names(hawk_df) %in% c("Month", "Year", "CaptureTime", "BandNumber", "Species", "Age", "Wing", "Weight", "Culmen", "Hallux", "Tail", "StandardTail", "Tarsus")]
hawk_df = replace_na(hawk_df, list(`Hallux` = mean(hawk_df$'Hallux', na.rm = TRUE)))
ch = subset(hawk_df, Species == "CH")
rt = subset(hawk_df, Species == "RT")
ss = subset(hawk_df, Species == "SS")
avg_ch = mean(ch$Hallux)
avg_rt = mean(rt$Hallux)
avg_ss = mean(ss$Hallux)
```

Species       | Hallux Avg
------------- | ----------
Cooper's      |  22.82      
Red-tailed    |  31.96       
Sharp-shinned |  15.10        

## Regression Model of the Capture Time vs the Age of Species of Hawks
```{r eval = TRUE, echo = FALSE, warning = FALSE}
setwd("C:/Users/jcuel/Videos/Bellevue/Summer 2021/Professor Williams/Projects/Project 1")
hawk_df = read.csv("Hawks.csv")
hawk_df = hawk_df %>% select(-X)
hawk_df = hawk_df[ , names(hawk_df) %in% c("Month", "Year", "CaptureTime", "BandNumber", "Species", "Age", "Wing", "Weight", "Culmen", "Hallux", "Tail", "StandardTail", "Tarsus")]
hawk_df = replace_na(hawk_df, list(`Hallux` = mean(hawk_df$'Hallux', na.rm = TRUE)))
tab_ct_sp_age = table(hawk_df$CaptureTime, hawk_df$Species, hawk_df$Age)
reg_mod1 = lm(Hallux ~ CaptureTime + Species + Age, data = hawk_df)
sum_ct_sp_age = summary(reg_mod1)
cfint_ct_sp_age = confint(reg_mod1)
plot(reg_mod1)
```
